### 2013-05-17 doc/pgas13/intro.tex
add to intro, include Grappa and Flat Combining sections

% The Grappa~\cite{Nelson:hotpar11-real} runtime system is optimized for fine-grained accesses.

@@

>>>>

@@

To honor the shared-memory abstraction, a common goal is for distributed shared data structures to adhere to a sequentially consistent memory model. Again, the simplest way to maintain a consistent global view of the data structure is to have a single global lock that is used to implementation mutual exclusion and enforce a serializable order over insertions. However, in a system with physically distributed memory, the cost of synchronizing in this way is even greater. In addition, with the massive amount of parallelism in a cluster of multiprocessors, the amount of contention on this shared data structure is higher than any shared-memory multiprocessor.

In many cases, algorithms do not require operations to be immediately consistent. Many optimizations for message-passing or PGAS-style distributed memory algorithms leverage this observation to minimize the need for fine-grained communication. However, they typically must make this visible to programmers, who now must take care to insert their own memory barriers where appropriate, or otherwise express the way in which consistency relaxation can occur.

The goal of this work is to see whether the concept of flat-combining can be used to make the cost of enforcing fine-grained synchronization acceptable.
To evaluate this possibility, we leverage the Grappa runtime~\cite{Nelson:hotpar11-real}, a PGAS-style runtime library optimized for fine-grained random access, which provides the ability to tolerate long latencies by efficiently switching between many lightweight threads.
With enough concurrent threads, it is actually possible to allow each thread to maintain sequentiality by blocking on accesses, but still have an opportunity to effectively combine many fine-grained synchronous operations for better performance.

% lay out the upcoming sections of the paper...
